They're Clutching up! Team Momentum in Round-Based Esports

### 2022-08-18 paper/paper.Rmd
reduce long eq 3
splitting up methodology and results

# Methodology and results

@@

Afterwards, I investigate post-streak win rates, building on my learnings from the constant probability assumption analysis. I agree with Miller and Sanjurjo's [-@miller2018] framework for streak selection bias, accounting for it in when analyzing momentum.

## Distribution of rounds played {#sec:analysis-1}

The general formula for the probability $P_E(r)$ that a best-of-$s$ series lasts $r$ rounds given constant probability $p$ of one team[^11] winning each round is

$$
q = 1 - p, s_1 = \frac{s_1 - 1}{2}, s_2 = \frac{s_1 + 1}{2}, s_2^{'} = 1 - s_2
$$

```{=tex}
\begin{equation}\protect\hypertarget{eq-series-length}{}{
P_E(r) = \frac{(r - 1)!}{s_1!(r - s_2)!}(p^{s_2}q^{s_2^{'}} + p^{s_2^{'}}q^{s_2}).
}\label{eq:series-length}\end{equation}
```
```{r}
#| label: cod_series_outcome_prop
.max_round <- 11
.cutoff <- 6
prob_of_series_lasting_n_games <- function(n, p = 0.5, max_round = .max_round) {
  s <- ceiling(max_round / 2)
  (factorial(n - 1) / (factorial(s - 1) * factorial(n - s))) * (p^s * (1 - p)^(n - s) + p^(n - s) * (1 - p)^s)
}

@@

  expected_round_streaks <- tibble(
    n_rounds = .cutoff:.max_round
  ) |> 
    mutate(
      series_prop = map_dbl(n_rounds, ~prob_of_series_lasting_n_games(.x, max_round = .max_round, p = !!p)),
      n_expected_series_streaks = map_dbl(n_rounds, ~expected_series_streaks_of_outcomes(.cutoff, .x - .cutoff))
    ) |> 
    transmute(
      n_rounds,
      series_prop,
      prop = series_prop / n_expected_series_streaks
    )

@@

Calculating the chi-square goodness of fit statistic

```{=tex}
\begin{equation}\protect\hypertarget{eq:series-length}{}{
\chi^2 = \sum^{11}_{i=6} \frac{(P_O(r) - P_E(r))^2}{P_E(r)}, r \in R = [6, 7, 8, 9, 10, 11].
}\label{eq:chi-squ-test}\end{equation}
```
as 16.0 (p-value of 0.0068), I can comfortably reject the constant probability null hypothesis for $p_0 = 0.5$, event at a confidence level of $\alpha = 0.01$.

@@

```{r}
#| label: cod-alternative-to-p=0.5-method-1
## p. 365 on https://math.mit.edu/classes/18.095/2016IAP/lec9/Sports_Mosteller1952_WorldSeries.pdf
theoretical_cod_series_length <- function(p) {
  6 * prob_of_series_lasting_n_games(n = 6, p = p, max_round = 11) +
    7 * prob_of_series_lasting_n_games(n = 7, p = p, max_round = 11) +
    8 * prob_of_series_lasting_n_games(n = 8, p = p, max_round = 11) +
    9 * prob_of_series_lasting_n_games(n = 9, p = p, max_round = 11) +
    10 * prob_of_series_lasting_n_games(n = 10, p = p, max_round = 11) +
    11 * prob_of_series_lasting_n_games(n = 11, p = p, max_round = 11)
}
cod_rounds_per_series <- cod_streaks_naive_res$series |>
  summarize(
    actual = sum(n_rounds * prop_actual),
    expected = sum(n_rounds * prop_expected)
  )

@@

maximize_cod_series_p <- function(p) {
  prob_of_series_lasting_n_games(n = 6, p = p, max_round = 11)^cod_n6adj *
    prob_of_series_lasting_n_games(n = 7, p = p, max_round = 11)^cod_n7adj *
    prob_of_series_lasting_n_games(n = 8, p = p, max_round = 11)^cod_n8adj *
    prob_of_series_lasting_n_games(n = 9, p = p, max_round = 11)^cod_n9adj *
    prob_of_series_lasting_n_games(n = 10, p = p, max_round = 11)^cod_n10adj *
    prob_of_series_lasting_n_games(n = 11, p = p, max_round = 11)^cod_n11adj
}

@@

Table \ref{tbl:mosteller-methods-results} shows the alternate values for the constant round win probability $p$ that I find when applying the three methods suggested by Mosteller [-@mosteller1952]. Each is approximately or equal to 0.575, and, when applying Equation \ref{eq:chi-squ-test}, each results in a $\chi^2$ value for which I cannot reject the constant probability null hypothesis.

```{=tex}
\begin{longtable}[]{@{}lrr@{}}
\caption{Alternate estimates of the constant probability \(p\) for winning a given round in a CoD SnD, applying the three methods suggested by Mosteller (1952), in addition to the naive \(p = 0.5\).}\label{tbl:mosteller-methods-results} \\
\toprule()
Method & \(p(r)\) & \(\chi^2\) (p-value) \\
\midrule()
\endfirsthead
\toprule()
Method & \(p(r)\) & \(\chi^2\) (p-value) \\
\midrule()
\endhead
0. Naive & 0.5000 & 16.0 (\textless=0.01) \\
1. Method of moments & 0.5725 & 3.6 (0.6) \\
2. Maximum likelihood & 0.5750 & 3.5 (0.62) \\
3. Minimum \(\chi^2\) & 0.5775 & 3.5 (0.62) \\
\bottomrule()
\end{longtable}
```
Table \ref{tbl:expected-series-lengths-alternative-ps} shows the new $P_E(r)$ when re-applying Equation \ref{eq:series-length} for the maximum likelihood estimate $p_2 = 0.575$.[^12] I observe that $P_E(r)$ is notably larger for $p_{2}$ when $r \in [6, 7]$, more closely matching $P_O(r)$. $p_{2}$ is also closer to $P_O(r)$ for $r \in [9, 10]$, although not for $r \in [8, 11]$.

[^12]: The method of moments and minimum $\chi^2$ estimates for $p$ are omitted simply because the results would be nearly identical to those for the maximum likelihood estimate of $p$ (since they are all $\approx 0.575$).

@@

## Momentum {#sec:analysis-2}

@@

Despite the implication that one can model team round win probability in CoD SnD as a constant, anecdotally many believe in momentum i.e. streaks not due to random chance. I now evaluate win frequency in rounds following a streak of length $k$. I account for the length of the series ($r$ rounds), modeling the expected round win probability after a streak of length $k$ as

```{=tex}
\begin{equation}\protect\hypertarget{eq:pwkr}{}{
  \hat{P}^{kr,\text{win}}_{E,\text{unadj.}} = \hat{P}(\text{win round | k-round winning streak, series lasting r rounds})=\{
    \begin{array}{cl}
    \frac{6 - k}{i - k}, & \text{team wins series} \\
    \frac{11 - 6 - k}{i - k}, & \text{team loses series}
    \end{array}.
  }\label{eq:pwkr}
\end{equation}
```
$\hat{P}^{kr,loss}_{E,\text{unadj.}} = \hat{P}(\text{lose round | k-round losing streak, series lasting r rounds})$ can be formulated symmetrically.

While one might think that $E[\hat{P}^{kr,\text{win}} - \hat{P}^{kr,\text{loss}}] = 0$, Miller and Sanjurjo [-@miller2018] suggest that, given a deterministic i.i.d. Bernoulli probability (Equation \ref{eq:pwkr}, this is not true due to streak-selection bias.[^13] As proved by Miller and Sanjurjo, TODO: Combine these two sentences

[^13]: Note that these probabilities are separate from those in Section \ref{sec:analysis-1}.

TODO: Descripe the Appendix E.1. I apply their formulation for "downward bias" to post-streak success to come up with estimates $\hat{P}^{kr,\text{win}}_{E,\text{adj.}}$, specifying $r$ as the number of trials.

Given that people typically perceive streaks as beginning after the third success (or failure) at minimum [@carlson2007], I focus on round win percentages in rounds immediately following streaks of three wins.[^14] Table \ref{tbl:cod-round-win-prop-after-k-wins} shows $\hat{P}^{kr,\text{win}}_O$ and $\hat{P}^{kr,\text{loss}}_O$ for $k = 3$ streaks respectively.

[^14]: Three happens to also be a reasonable number for series that lasts at maximum 11 rounds.

>>>>

# Methodology

## Distribution of rounds played {#sec:method-rounds-played}

The general formula for the probability $P_E(r)$ that a best-of-$s$ series ($s = 11$ for CoD SnD) lasts $r$ rounds given constant probability $p$ of one team[^11] winning each round is

$$
m = \frac{s + 1}{2}
$$

```{=tex}
\begin{equation}\protect\hypertarget{eq:series-length}{}{
P_E(r) = \frac{(r - 1)!}{(m - 1)!(r - s)!}(p^{m}(1 - p)^{r - m} + p^{r - m}(1 - p)^m).
}\label{eq:series-length}\end{equation}
```
The chi-square goodness of fit statistic can be used to test the constant found probability hypothesis for $0 \leq p \leq 1$.

```{=tex}
\begin{equation}\protect\hypertarget{eq:chi-squ}{}{
\chi^2 = \sum^R \frac{(P_O(r) - P_E(r))^2}{P_E(r)}
}\label{eq:chi-squ}\end{equation}
```
For CoD SnD, $r \in R [6, 7, 8, 9, 10, 11]$.

## Momentum {#sec:method-momentum}

Without accounting for the streak selection bias described by Miller and Sanjurjo [-@miller2018], the round win probability after a streak of $k$ round wins for a best-of-$s$ series ($s = 11$ for CoD SnD), accounting for the series lasting $r$ rounds known in hindsight, is

$$
m = \frac{s + 1}{2}
$$

```{=tex}
\begin{equation}\protect\hypertarget{eq:pwkr}{}{
  p^{w|kr} = p(\text{win | k, r})=\{
    \begin{array}{cl}
    \frac{m - k}{i - k}, & \text{team wins series} \\
    \frac{s - m - k}{i - k}, & \text{team loses series}
    \end{array}.
  }
\label{eq:pwkr}
\end{equation}
```
The "hindsight" round loss probability after a streak of $k$ round losses for a series lasting $r$ rounds, $p^{l|kr}$, can be formulated symmetrically.[^12]

[^12]: Note that these probabilities are separate from the constant round probabilities in Sections \ref{sec:method-rounds-played} and \ref{sec:results-rounds-played}.

However, as shown by Miller and Sanjurjo, we should expect the observed proportion of rounds wins after a streak $k$ rounds wins for a series lasting $r$ rounds, $\hat{P}^{w|kr}$, to be strictly less than $p^{w|kr}$.[^13]

[^13]: See Appendix A of Miller and Sanjurjo [-@miller2018] for the proof.

Miller and Sanjurjo [-@miller2018] note that there does not exists a closed form representation for $E[\hat{P}_k(\mathbf{X}) | I_k(\mathbf{X}) \neq \emptyset]$ ($\hat{P}^{w|kr}$ for CoD SnD) for $k > 1$. Figure \ref{fig:npk} shows this.[^14]

[^14]: I've adapted the code from Vafa [-@vafa2017].

![The expected value](images/npk_replicated.png)

The expected value of the proportion of successes on trials that immediately follow k consecutive successes, Pˆk(X), as a function of the total number of trials n,for different values of k and probabilities of success p,

# Results

@@

## Distribution of rounds played {#sec:results-rounds-played}

Using Equation \ref{eq:chi-squ}, $\chi^2 = 16.0$ (p-value of 0.0068) for a naive constant round win probability $p_0 = 0.5$. Thus, I can comfortably reject the constant probability null hypothesis for $p_0 = 0.5$, even at a confidence level of $\alpha = 0.01$. Afterwards, I investigate post-streak win rates, building on my learnings from the constant probability assumption analysis. I agree with Miller and Sanjurjo's [-@miller2018] framework for streak selection bias, accounting for it in when analyzing momentum.

```{r}
#| label: cod_series_outcome_prop
.max_round <- 11
.cutoff <- 6
prob_of_series_lasting_r_rounds <- function(r, p = 0.5, s = .max_round) {
  m <- as.integer((s + 1) / 2)
  (factorial(r - 1) / (factorial(m - 1) * factorial(r - s))) * (p^m * (1 - p)^(r - m) + p^(r - m) * (1 - p)^m)
}

@@

  expected_round_streaks <- tibble(
    n_rounds = .cutoff:.max_round
  ) |> 
    mutate(
      series_prop = map_dbl(n_rounds, ~prob_of_series_lasting_r_rounds(.x, max_round = .max_round, p = !!p)),
      n_expected_series_streaks = map_dbl(n_rounds, ~expected_series_streaks_of_outcomes(.cutoff, .x - .cutoff))
    ) |> 
    transmute(
      n_rounds,
      series_prop,
      prop = series_prop / n_expected_series_streaks
    )

@@

@@

```{r}
#| label: cod-alternative-to-p=0.5-method-1
## p. 365 on https://math.mit.edu/classes/18.095/2016IAP/lec9/Sports_Mosteller1952_WorldSeries.pdf
theoretical_cod_series_length <- function(p) {
  6 * prob_of_series_lasting_r_rounds(n = 6, p = p, max_round = 11) +
    7 * prob_of_series_lasting_r_rounds(n = 7, p = p, max_round = 11) +
    8 * prob_of_series_lasting_r_rounds(n = 8, p = p, max_round = 11) +
    9 * prob_of_series_lasting_r_rounds(n = 9, p = p, max_round = 11) +
    10 * prob_of_series_lasting_r_rounds(n = 10, p = p, max_round = 11) +
    11 * prob_of_series_lasting_r_rounds(n = 11, p = p, max_round = 11)
}
cod_rounds_per_series <- cod_streaks_naive_res$series |>
  summarize(
    actual = sum(n_rounds * prop_actual),
    expected = sum(n_rounds * prop_expected)
  )

@@

maximize_cod_series_p <- function(p) {
  prob_of_series_lasting_r_rounds(n = 6, p = p, max_round = 11)^cod_n6adj *
    prob_of_series_lasting_r_rounds(n = 7, p = p, max_round = 11)^cod_n7adj *
    prob_of_series_lasting_r_rounds(n = 8, p = p, max_round = 11)^cod_n8adj *
    prob_of_series_lasting_r_rounds(n = 9, p = p, max_round = 11)^cod_n9adj *
    prob_of_series_lasting_r_rounds(n = 10, p = p, max_round = 11)^cod_n10adj *
    prob_of_series_lasting_r_rounds(n = 11, p = p, max_round = 11)^cod_n11adj
}

@@

Table \ref{tbl:mosteller-methods-results} shows the alternate values for the constant round win probability $p$ that I find when applying the three methods suggested by Mosteller [-@mosteller1952]. Each is approximately or equal to 0.575, and, when applying Equation \ref{eq:chi-squ}, each results in a $\chi^2$ value for which I cannot reject the constant probability null hypothesis.

```{=tex}
\begin{longtable}[]{@{}lrr@{}}
\caption{Alternate estimates of the constant probability \(p\) for winning a given round in a CoD SnD, applying the three methods suggested by Mosteller (1952), in addition to the naive \(p = 0.5\).}\label{tbl:mosteller-methods-results} \\
\toprule()
Method & \(p(r)\) & \(\chi^2\) (p-value) \\
\midrule()
\endfirsthead
\toprule()
Method & \(p(r)\) & \(\chi^2\) (p-value) \\
\midrule()
\endhead
0. Naive & 0.5000 & 16.0 (\textless=0.01) \\
1. Method of moments & 0.5725 & 3.6 (0.6) \\
2. Maximum likelihood & 0.5750 & 3.5 (0.62) \\
3. Minimum \(\chi^2\) & 0.5775 & 3.5 (0.62) \\
\bottomrule()
\end{longtable}
```
Table \ref{tbl:expected-series-lengths-alternative-ps} shows the new $P_E(r)$ when re-applying Equation \ref{eq:series-length} for the maximum likelihood estimate $p_2 = 0.575$.[^15] I observe that $P_E(r)$ is notably larger for $p_{2}$ when $r \in [6, 7]$, more closely matching $P_O(r)$. $p_{2}$ is also closer to $P_O(r)$ for $r \in [9, 10]$, although not for $r \in [8, 11]$.

[^15]: The method of moments and minimum $\chi^2$ estimates for $p$ are omitted simply because the results would be nearly identical to those for the maximum likelihood estimate of $p$ (since they are all $\approx 0.575$).

@@

## Momentum {#sec:results-momentum}

@@

Given that people typically perceive streaks as beginning after the third success (or failure) at minimum [@carlson2007], I focus on round win percentages in rounds immediately following streaks of three wins.[^16] Table \ref{tbl:cod-round-win-prop-after-k-wins} shows $\hat{P}^{kr,\text{win}}_O$ and $\hat{P}^{kr,\text{loss}}_O$ for $k = 3$ streaks respectively.

[^16]: Three happens to also be a reasonable number for series that lasts at maximum 11 rounds.

### 2022-08-18 paper/paper.tex
splitting up methodology and results

\hypertarget{methodology-and-results}{%
\section{Methodology and results}\label{methodology-and-results}}

@@

Afterwards, I investigate post-streak win rates, building on my
learnings from the constant probability assumption analysis. I agree
with Miller and Sanjurjo's (2018) framework for streak selection bias,
accounting for it in when analyzing momentum.

\hypertarget{sec:analysis-1}{%
\subsection{Distribution of rounds played}\label{sec:analysis-1}}

The general formula for the probability \(P_E(r)\) that a best-of-\(s\)
series lasts \(r\) rounds given constant probability \(p\) of one
team\footnote{If \(p > 0.5\), then one might say that this team is the
  better team (known in hindsight).} winning each round is

\[
q = 1 - p, s_1 = \frac{s_1 - 1}{2}, s_2 = \frac{s_1 + 1}{2}, s_2^{'} = 1 - s_2
\]

\begin{equation}\protect\hypertarget{eq-series-length}{}{
P_E(r) = \frac{(r - 1)!}{s_1!(r - s_2)!}(p^{s_2}q^{s_2^{'}} + p^{s_2^{'}}q^{s_2}).
}\label{eq:series-length}\end{equation}

@@

Calculating the chi-square goodness of fit statistic

\begin{equation}\protect\hypertarget{eq:series-length}{}{
\chi^2 = \sum^{11}_{i=6} \frac{(P_O(r) - P_E(r))^2}{P_E(r)}, r \in R = [6, 7, 8, 9, 10, 11].
}\label{eq:chi-squ-test}\end{equation}

as 16.0 (p-value of 0.0068), I can comfortably reject the constant
probability null hypothesis for \(p_0 = 0.5\), event at a confidence
level of \(\alpha = 0.01\).

Table \ref{tbl:mosteller-methods-results} shows the alternate values for
the constant round win probability \(p\) that I find when applying the
three methods suggested by Mosteller (1952). Each is approximately or
equal to 0.575, and, when applying Equation \ref{eq:chi-squ-test}, each
results in a \(\chi^2\) value for which I cannot reject the constant
probability null hypothesis.

@@

\hypertarget{sec:analysis-2}{%
\subsection{Momentum}\label{sec:analysis-2}}

Despite the implication that one can model team round win probability in
CoD SnD as a constant, anecdotally many believe in momentum i.e.~streaks
not due to random chance. I now evaluate win frequency in rounds
following a streak of length \(k\). I account for the length of the
series (\(r\) rounds), modeling the expected round win probability after
a streak of length \(k\) as

\begin{equation}\protect\hypertarget{eq:pwkr}{}{
  \hat{P}^{kr,\text{win}}_{E,\text{unadj.}} = \hat{P}_{E,\text{unadj.}}(\text{win | k, r})=\{
    \begin{array}{cl}
    \frac{6 - k}{i - k}, & \text{team wins series} \\
    \frac{11 - 6 - k}{i - k}, & \text{team loses series}
    \end{array}.
  }\label{eq:pwkr}
\end{equation}

\(\hat{P}^{kr,loss}_{E,\text{unadj.}}\) can be formulated symmetrically.

While one might think that
\(E[\hat{P}^{kr,\text{win}} - \hat{P}^{kr,\text{loss}}] = 0\), Miller
and Sanjurjo (2018) suggest that, given a deterministic i.i.d. Bernoulli
probability (Equation \ref{eq:pwkr}, this is not true due to
streak-selection bias.\footnote{Note that these probabilities are
  separate from those in Section \ref{sec:analysis-1}.} As proved by
Miller and Sanjurjo, TODO: Combine these two sentences

TODO: Descripe the Appendix E.1. I apply their formulation for
``downward bias'' to post-streak success to come up with estimates
\(\hat{P}^{kr,\text{win}}_{E,\text{adj.}}\), specifying \(r\) as the
number of trials.

@@

\begin{tabbing}
\begin{longtable}{rrrrrrr}
\caption{First, observed proportion of wins ($\hat{P}^{kr,\text{win}}_O$) and expected proportion of wins, both un-adjusted ($\hat{P}^{kr,\text{win}}_{E,\text{unadj.}}$) and adjusted ($\hat{P}^{kr,\text{win}}_{E,\text{adj.}}$), after 3-round win streak, for a series lasting $r$ rounds. Second, observed proportion of losses ($\hat{P}^{kr,\text{loss}}_O$) and expected proportion of losses, both un-adjusted ($\hat{P}^{kr,\text{loss}}_{E,\text{unadj.}}$) and adjusted ($\hat{P}^{kr,\text{loss}}_{E,\text{adj.}}$), after 3-round loss streak}\label{tbl:cod-round-win-prop-after-k-wins} \\
\toprule
\multicolumn{7}{c}{\text{3 round win streak}} \\
 &  &  & \multicolumn{2}{c}{\text{Unadjusted}} & \multicolumn{2}{c}{\text{Adjusted}} \\ 
\cmidrule(lr){4-5} \cmidrule(lr){6-7}
$r$ & $N^{kr,\text{win}}_O$ & $\hat{P}^{kr,\text{win}}_O$ & $\hat{P}^{kr,\text{win}}_{E,\text{unadj.}}$ &  $p$-value & $\hat{P}^{kr,\text{win}}_{E,\text{adj.}}$ & $p$-value \\ 
\midrule
7 & 209 & $75\%$ & $75\%$ & 0.58 & $61\%$ & <=0.01 \\ 
8 & 209 & $62\%$ & $60\%$ & 0.28 & $45\%$ & <=0.01 \\ 
9 & 193 & $52\%$ & $50\%$ & 0.33 & $35\%$ & <=0.01 \\ 
10 & 151 & $44\%$ & $43\%$ & 0.45 & $29\%$ & <=0.01 \\ 
11 & 150 & $40\%$ & $38\%$ & 0.29 & $25\%$ & <=0.01 \\ 
\toprule
\multicolumn{7}{c}{\text{3 round loss streak}} \\
 &  &  & \multicolumn{2}{c}{\text{Unadjusted}} & \multicolumn{2}{c}{\text{Adjusted}} \\ 
\cmidrule(lr){4-5} \cmidrule(lr){6-7}
$r$ & $N^{kr,\text{loss}}_O$ & $\hat{P}^{kr,\text{loss}}_O$ & $\hat{P}^{kr,\text{loss}}_{E,\text{unadj.}}$ &  $p$-value & $\hat{P}^{kr,\text{loss}}_{E,\text{adj.}}$ & $p$-value \\ 
\midrule
10 & 60 & $13\%$ & $14\%$ & 0.64 & $10\%$ & 0.26 \\ 
11 & 129 & $24\%$ & $25\%$ & 0.63 & $16\%$ & <=0.01 \\ 
\bottomrule
\end{longtable}
\end{tabbing}

>>>>

\hypertarget{methodology}{%
\section{Methodology}\label{methodology}}

\hypertarget{sec:method-rounds-played}{%
\subsection{Distribution of rounds
played}\label{sec:method-rounds-played}}

The general formula for the probability \(P_E(r)\) that a best-of-\(s\)
series (\(s = 11\) for CoD SnD) lasts \(r\) rounds given constant
probability \(p\) of one team\footnote{If \(p > 0.5\), then one might
  say that this team is the better team (known in hindsight).} winning
each round is

\[
m = \frac{s + 1}{2}
\]

\begin{equation}\protect\hypertarget{eq:series-length}{}{
P_E(r) = \frac{(r - 1)!}{(m - 1)!(r - s)!}(p^{m}(1 - p)^{r - m} + p^{r - m}(1 - p)^m).
}\label{eq:series-length}\end{equation}

The chi-square goodness of fit statistic can be used to test the
constant found probability hypothesis for \(0 \leq p \leq 1\).

\begin{equation}\protect\hypertarget{eq:chi-squ}{}{
\chi^2 = \sum^R \frac{(P_O(r) - P_E(r))^2}{P_E(r)}
}\label{eq:chi-squ}\end{equation}

For CoD SnD, \(r \in R [6, 7, 8, 9, 10, 11]\).

\hypertarget{sec:method-momentum}{%
\subsection{Momentum}\label{sec:method-momentum}}

Without accounting for the streak selection bias described by Miller and
Sanjurjo (2018), the round win probability after a streak of \(k\) round
wins for a best-of-\(s\) series (\(s = 11\) for CoD SnD), accounting for
the series lasting \(r\) rounds known in hindsight, is

\[
m = \frac{s + 1}{2}
\]

\begin{equation}\protect\hypertarget{eq:pwkr}{}{
  p^{w|kr} = p(\text{win | k, r})=\{
    \begin{array}{cl}
    \frac{m - k}{i - k}, & \text{team wins series} \\
    \frac{s - m - k}{i - k}, & \text{team loses series}
    \end{array}.
  }
\label{eq:pwkr}
\end{equation}

The ``hindsight'' round loss probability after a streak of \(k\) round
losses for a series lasting \(r\) rounds, \(p^{l|kr}\), can be
formulated symmetrically.\footnote{Note that these probabilities are
  separate from the constant round probabilities in Sections
  \ref{sec:method-rounds-played} and \ref{sec:results-rounds-played}.}

However, as shown by Miller and Sanjurjo, we should expect the observed
proportion of rounds wins after a streak \(k\) rounds wins for a series
lasting \(r\) rounds, \(\hat{P}_{w|kr}\), to be strictly less than
\(p^{w|kr}\).

Miller and Sanjurjo (2018) note that there does not exists a closed form
representation for
\(E[\hat{P}_k(\mathbf{X}) | I_k(\mathbf{X}) \neq \emptyset]\) for
\(k > 1\).

\hypertarget{results}{%
\section{Results}\label{results}}

@@

\hypertarget{sec:results-rounds-played}{%
\subsection{Distribution of rounds
played}\label{sec:results-rounds-played}}

Using Equation \ref{eq:chi-squ}, \(\chi^2 = 16.0\) (p-value of 0.0068)
for a naive constant round win probability \(p_0 = 0.5\). Thus, I can
comfortably reject the constant probability null hypothesis for
\(p_0 = 0.5\), even at a confidence level of \(\alpha = 0.01\).
Afterwards, I investigate post-streak win rates, building on my
learnings from the constant probability assumption analysis. I agree
with Miller and Sanjurjo's (2018) framework for streak selection bias,
accounting for it in when analyzing momentum.

@@

Table \ref{tbl:mosteller-methods-results} shows the alternate values for
the constant round win probability \(p\) that I find when applying the
three methods suggested by Mosteller (1952). Each is approximately or
equal to 0.575, and, when applying Equation \ref{eq:chi-squ}, each
results in a \(\chi^2\) value for which I cannot reject the constant
probability null hypothesis.

@@

\hypertarget{sec:results-momentum}{%
\subsection{Momentum}\label{sec:results-momentum}}

@@

\begin{longtable}{rrrrrrr}
\caption{First, observed proportion of wins ($\hat{P}^{kr,\text{win}}_O$) and expected proportion of wins, both un-adjusted ($\hat{P}^{kr,\text{win}}_{E,\text{unadj.}}$) and adjusted ($\hat{P}^{kr,\text{win}}_{E,\text{adj.}}$), after 3-round win streak, for a series lasting $r$ rounds. Second, observed proportion of losses ($\hat{P}^{kr,\text{loss}}_O$) and expected proportion of losses, both un-adjusted ($\hat{P}^{kr,\text{loss}}_{E,\text{unadj.}}$) and adjusted ($\hat{P}^{kr,\text{loss}}_{E,\text{adj.}}$), after 3-round loss streak}\label{tbl:cod-round-win-prop-after-k-wins} \\
\toprule
\multicolumn{7}{c}{\text{3 round win streak}} \\
 &  &  & \multicolumn{2}{c}{\text{Unadjusted}} & \multicolumn{2}{c}{\text{Adjusted}} \\ 
\cmidrule(lr){4-5} \cmidrule(lr){6-7}
$r$ & $N^{kr,\text{win}}_O$ & $\hat{P}^{kr,\text{win}}_O$ & $\hat{P}^{kr,\text{win}}_{E,\text{unadj.}}$ &  $p$-value & $\hat{P}^{kr,\text{win}}_{E,\text{adj.}}$ & $p$-value \\ 
\midrule
7 & 209 & $75\%$ & $75\%$ & 0.58 & $61\%$ & <=0.01 \\ 
8 & 209 & $62\%$ & $60\%$ & 0.28 & $45\%$ & <=0.01 \\ 
9 & 193 & $52\%$ & $50\%$ & 0.33 & $35\%$ & <=0.01 \\ 
10 & 151 & $44\%$ & $43\%$ & 0.45 & $29\%$ & <=0.01 \\ 
11 & 150 & $40\%$ & $38\%$ & 0.29 & $25\%$ & <=0.01 \\ 
\toprule
\multicolumn{7}{c}{\text{3 round loss streak}} \\
 &  &  & \multicolumn{2}{c}{\text{Unadjusted}} & \multicolumn{2}{c}{\text{Adjusted}} \\ 
\cmidrule(lr){4-5} \cmidrule(lr){6-7}
$r$ & $N^{kr,\text{loss}}_O$ & $\hat{P}^{kr,\text{loss}}_O$ & $\hat{P}^{kr,\text{loss}}_{E,\text{unadj.}}$ &  $p$-value & $\hat{P}^{kr,\text{loss}}_{E,\text{adj.}}$ & $p$-value \\ 
\midrule
10 & 60 & $13\%$ & $14\%$ & 0.64 & $10\%$ & 0.26 \\ 
11 & 129 & $24\%$ & $25\%$ & 0.63 & $16\%$ & <=0.01 \\ 
\bottomrule
\end{longtable}
